Comb associative operads

### 2018-08-28 Introduction.tex
Correction coquilles intro

To explore this vast research area, we select to follow in this paper
the following directions. First, we consider the very general set of the
quotients of $\Mag$ seen as a operad in the category of vector spaces.
We show that these operads form a lattice, wherein its partial order
relation is defined from the existence of operad morphisms
(Theorem~\ref{thm:lattice_structure_of_QMag}). We also provide a
Grassmann formula analog relating the Hilbert series of the operads of
the lattice together with their infima and suprema
(Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}).
Besides, we focus on a special kind of quotients of $\Mag$, denoted by
$\CAs{\gamma}$, defined by equating the left and right comb binary trees
of a fixed degree $\gamma \geq 1$. These operads are called comb
associative operads. For instance, $\CAs{3}$ is the operad describing
the category of the algebras equipped with a binary product $\Product$
subjected to the relation
\begin{equation}
    \left(\left(x_1 \Product x_2\right) \Product x_3\right) \Product x_4
    =
    x_1 \Product \left(x_2 \Product \left(x_3 \Product x_4\right)\right).
\end{equation}
This is a kind of associativity law in higher degree $\gamma = 3$ so
that $\CAs{3}$ can be seen as a generalization of $\As$ (which is equal
to $\CAs{2}$). We first provide general results about the operads
$\CAs{\gamma}$. In particular, we show that these operads form a lattice
contained in the lattice of quotients of $\Mag$ aforementioned
(Theorems~\ref{thm:lattice_CAs} and~\ref{thm:inclusion_lattice_CAs}). We
focus in particular on the study of $\CAs{3}$. Observe that the
congruence of this operad is generated by an equivalence class of trees
which is not an interval for the Tamari order. As preliminary computer
experiments show, $\CAs{3}$ has oscillating first dimensions
(see~\eqref{equ:dimensions_CAs_3}), what is rather unusual among all
known operads. We provide an orientation of the space of relations of
$\CAs{3}$ (Theorem~\ref{thm:convergent_rewrite_rule_CAs_3}), a
description of a basis of the operad, and prove that its Hilbert series
is rational.
%
\Todo{S: et parler de la réalisation si Christophe l'ajoute ?}
%
For all these, we use rewrite systems on trees~\cite{BN98} and the
Buchberger algorithm for operads~\cite{DK10}. Finally, we continue the
investigation of the quotients of $\Mag$ by regarding the quotients of
$\Mag$ obtained by equating two trees of degree $3$. This leads to ten
quotient operads of $\Mag$. We provide for some of these combinatorial
realizations, mostly in terms of integer compositions.
\medbreak

>>>>

To explore this vast research area, we select to follow in this paper
the following directions. First, we consider the very general set of the
quotients of $\Mag$ seen as a operad in the category of vector spaces.
We show that these operads form a lattice, wherein its partial order
relation is defined from the existence of operad morphisms
(Theorem~\ref{thm:lattice_structure_of_QMag}). We also provide a
Grassmann formula analog relating the Hilbert series of the operads of
the lattice together with their infima and suprema
(Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}).
Besides, we focus on a special kind of quotients of $\Mag$, denoted by
$\CAs{\gamma}$, defined by equating the left and right comb binary trees
of a fixed degree $\gamma \geq 1$. These operads are called comb
associative operads. For instance, $\CAs{3}$ is the operad describing
the category of the algebras equipped with a binary product $\Product$
subjected to the relation
\begin{equation}
    \left(\left(x_1 \Product x_2\right) \Product x_3\right) \Product x_4
    =
    x_1 \Product \left(x_2 \Product \left(x_3 \Product x_4\right)\right).
\end{equation}
This is a kind of associativity law in higher degree $\gamma = 3$ so
that $\CAs{3}$ can be seen as a generalization of $\As$ (which is equal
to $\CAs{2}$). We first provide general results about the operads
$\CAs{\gamma}$. In particular, we show that these operads form a lattice
contained in the lattice of quotients of $\Mag$ aforementioned
(Theorems~\ref{thm:lattice_CAs} and~\ref{thm:inclusion_lattice_CAs}). We
focus in particular on the study of $\CAs{3}$. Observe that the
congruence of this operad is generated by an equivalence class of trees
which is not an interval for the Tamari order. As preliminary computer
experiments show, $\CAs{3}$ has oscillating first dimensions
(see~\eqref{equ:dimensions_CAs_3}), what is rather unusual among all
known operads. We provide a convergent orientation of the space of
relations of $\CAs{3}$ (Theorem~\ref{thm:convergent_rewrite_rule_CAs_3}),
a description of a basis of the operad, and prove that its Hilbert series
is rational.
%
\Todo{S: et parler de la réalisation si Christophe l'ajoute ?}
%
For all these, we use rewrite systems on trees~\cite{BN98} and the
Buchberger algorithm for operads~\cite{DK10}. Finally, we continue the
investigation of the quotients of $\Mag$ by regarding the quotients of
$\Mag$ obtained by equating two trees of degree $3$. This leads to ten
quotient operads of $\Mag$. We provide for some of these combinatorial
realizations, mostly in terms of integer compositions.
\medbreak
